EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-28 03_edipack2.tex
Huge iteration
- Igor passed through a major part of the manuscript
- Samuele edited some parts of the e-ph section (response to Massimo)
- Adriano added author contributions
- Gabriele answered some comments by Igor and Lorenzo on the iRDM

--------

Co-authored-by: Igor Krivenko <iskrivenko@proton.me>
Co-authored-by: Lorenzo Crippa <crippa.lorenzo@gmail.com>
Co-authored-by: Samuele Giuli <samuele.giuli@gmail.com>
Co-authored-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\section{Implementation}\label{SecEDIpack}
In this section we introduce the generic quantum impurity system and
present a detailed overview of the \NAME library implementation.

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a quantum impurity problem described by the Hamiltonian:
$$
\hat{H} = \hat{H}_\mathrm{imp} + \hat{H}_\mathrm{bath} + \hat{H}_\mathrm{hyb} + \hat{H}_\mathrm{ph} + \hat{H}_\mathrm{e-ph}
$$
which characterizes a multi-orbital quantum impurity
coupled to an electronic bath and local phonons (e.g., Holstein
modes). For now, we assume no specific symmetries.

\subsubsection{Impurity Hamiltonian}
The impurity Hamiltonian is split into a quadratic and an 
interacting part:
\begin{equation}\label{Himp}
  \hat{H}_\mathrm{imp}  = \hat{H}^0_\mathrm{imp} + \hat{H}^\mathrm{int}_\mathrm{imp}.
\end{equation}
The first term has the form: 
\begin{equation}\label{H0imp}
  \hat{H}^0_\mathrm{imp}  =
  \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}d_{\b\sigma'}
\end{equation}
and represents the quadratic part, with $d_{\a\sigma}$ ($d^+_{\a\sigma}$) are the annihilation (creation)
operators for impurity electrons in 
orbital $\alpha=1,\dots,N_\a$, with $N_\a$ the number of orbitals,
and spin $\sigma=\up,\dw$. 
The internal structure of the non-interacting part is captured by the 
matrix $h^{0}_{\a\b\sigma\sigma'}$, which can include 
orbital-dependent on-site energies or amplitudes, spin-orbit coupling or other single-particle terms.

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can in principle contain any 
set of two-body operators: 
\begin{equation}
\hat{H}^\mathrm{int}_\mathrm{imp} = \frac{1}{2} \sum_{ijkl} c^{\dagger}_{i}c^{\dagger}_{j}U_{ijkl}c_{l}c_{k}.\label{HintUmat}
\end{equation}
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction~\cite{Georges2013ACMP}:
\begin{equation}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^{+}_{\a\uparrow}d_{\a\downarrow}d^{+}_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^{+}_{\a\uparrow}d^{+}_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow}\\
\end{split}
\label{Hint}
\end{equation}
where we introduced the occupation number operator   $n_{\alpha\sigma}=d^{+}_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Georges2013ACMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and
the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ACMP}. Different choices, preserving part
of the combined symmetry group, can be made for other number of orbitals~\cite{Georges2013ACMP}.

\subsubsection{Bath and Hybrization}
The coupling between the impurity and the bath is described by
\begin{equation}\label{Hbath}
  \begin{split}
    \hat{H}_\mathrm{bath} &=
    \sum_p\sum_{\a\b\sigma\sigma'}h^p_{\a\b\sigma\sigma'}a^{+}_{p\a\sigma}a_{p\b\sigma'}\\
    %
    \hat{H}_\mathrm{hyb} &= \sum_p\sum_{\a\b\sigma\sigma'}V^p_{\a\b\sigma\sigma'}d^{+}_{\a\sigma}a_{p\b\sigma'}+H.c. \\
\end{split}
\end{equation}
where $p=1,\dots,N_\mathrm{bath}$ indexes the
bath elements. The operators $a_{p\alpha\sigma}$
($a^+_{p\alpha\sigma}$) correspond to the destruction (creation) of
bath electrons with index $p$, orbital $\alpha$ and spin $\sigma$.
Any bath element can be composed of several electronic levels
according to the chosen bath topology (see \secu{sSecBath}). The  properties of each bath element are described by the
matrix $h^p_{\a\b\sigma\sigma'}$ while the amplitude
$V^p_{\a\b\s\sigma'}$ desribes the coupling with the impurity.

\subsubsection{Electron-Phonon coupling}
Finally, we include a local electron-phonon coupling on the impurity site
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_m \omega_{0m} b_m^+b_m   \\
    % 
    \hat{H}_\mathrm{e-ph} &= \sum_m g_m[\hat{O}_m - A_m](b_m+b_m^+) \\
\end{split}
\end{equation}
where $m=1,\dots,M$ indexes the number of local phonon modes, and $b_m$
($b_m^+$) are the destruction (creation) operators for a phonon with
frequency $\omega_{0m}$. Each phonon mode is coupled to a bilinear fermionic operator $\hat{O}_m$ on the impurity, while $A_m$ is a constant acting as a displacement field.  %Indeed, if the expectation value of the bosonic displacement operator $\langle b_q^\dagger + b_q\rangle$ is non-zero, a change of variables from the bosons $b^{(\dagger)}_q$ leads to new effective bosons $\tilde{b}^{(\dagger)}_q = b^{(\dagger)}_q- A_q$ for which the displacement operator has a smaller expectation value (ideally zero if we can impose $A_m = \langle \hat{O}_m\rangle$). 
For single-orbital impurities, the relevant case is $O_1 = n_d = \sum_{\sigma} d^{\dagger}_{\sigma}d_{\sigma}$, which defines the Holstein model\cite{Holstein1959APNY}. For two-orbital models we included also the two Jahn-Teller modes\cite{Capone2010ACMP}, for which $O_2 = n_1 - n_2$\cite{Scazzola2023PRB} and $O_3 = \sum_{\sigma} (d^{\dagger}_{1\sigma}d_{2\sigma}+ h.c.)$. For each phonon mode, one has to introduce a cut-off in the phonon number which has to be chosen according to the value electron-phonon coupling\cite{Capone2003PRL,Capone2006PRB} and its interplay with the electron-electron interaction\cite{Capone2004PRL_2,Sangiovanni2005PRB,Paci2005PRL,Sangiovanni2006PRL,Sangiovanni2006PRB_ph,Paci2006PRB}. The larger the number of excited phonons in the groundstate, the larger the cut-off. For practical calculations around 20 phonons are sufficient to achieve converged results except for the deep polaronic regime\cite{Capone2003PRL}.
Choosing properly the shift $A_m$ one can reduce the number of excited phonons in the groundstate and consequently the required cut-off. 
Although feasible, dealing with more than one phonon mode becomes quickly computationally very demanding, thus in the rest of the this work we will consider only one mode.

The phonon frequency is passed to the \NAME program via the variable { \tt W0\_PH } to be read from the input file. The displacement field acting on the phonon is passed through the variable {\tt A\_PH} from the input file. For every phonon mode, the $g_m$ coupling can be written as a matrix $g_{\alpha \beta}$ which can be passed in two different ways. If the matrix is diagonal ($g_{\alpha \beta} = g_\alpha \delta_{\alpha \beta}$), from the input file one can simply write in order the values $g_\alpha$ in the variable {\tt G\_PH }. For any general shape one can pass to the input file a {\tt GPHfile } variable containing the name of the file storing the $g_{\alpha \beta}$ array in Fortran style.
If {\tt GPHfile=NONE} the matrix $g_{\alpha \beta}$ is read from the variable {\tt G\_PH}, else it is read from the {\tt GPHfile}.

\subsubsection{System setup}
We consider a quantum impurity system composed by a single multi-orbital impurity, a bath discretized into a number $N_\mathrm{bath}$ of {\it elements} possibly endowed with an internal structure, and a number of available phonons. 
The total size of the system is determined by three contributions: i) the number of impurity orbitals $N_\a$, ii) the total count of electronic bath levels $N_\mathrm{b}$, and iii) by maximum  number of allowed phonons $N_\mathrm{ph}$, introducing a cut-off to the unbounded dimensions of the local phonon Hilbert space.
The number of bath degrees of freedom $N_\mathrm{b}$ is a
function of the bath topology and of $N_\mathrm{bath}$, which is determined internally by \NAME.
For instance, for a simple case of a purely electronic problem, each bath
element corresponds to an independent electronic level coupled to the
impurity, thus $N_\mathrm{b}\equiv N_\mathrm{bath}$. 
The total number of electronic levels in the quantum impurity systems is indicated with $N_\mathrm{s}$.

The setup of the quantum impurity problem is implemented in {\tt
  ED\_SETUP} through the input variables {\tt Nspin}$=N_\sigma$,
{\tt Norb}$=N_\a$ and {\tt Nbath}$=N_\mathrm{bath}$ globally defined {\tt
  ED\_INPUT\_VARS}. The {\tt Ns}$=N_\mathrm{s}$ variable, evaluated in
{\tt ed\_setup\_dimensions}, corresponds to the total number of
electronic levels. 
The local non-interacting Hamiltonian
$h^0_{\a\b\s\s'}$ is specified using the function {\tt ed\_set\_hloc} in {\tt ED\_AUX\_FUNX}.
The setup of the bath matrices $h^p_{\a\b\s\s'}$
requires a more involved procedure which will be illustrated in
\secu{sSecBath}.

@@

The electronic part of the Fock state $\ket{\vec{n}}$ is represented
as a binary string of length $2N_\mathrm{s}$. Thus, any such state can be encoded
in a computer using a sequence of $2N_\mathrm{s}$ bits, or equivalently, as an integer $I=0,\dots 2^{2N_\mathrm{s}}-1$ such that $\ket{\vec{n}}=\ket{I}$.  
The electornic destruction and creation operators, $c_{a\sigma}$ and $c^+_{a\sigma}$ 
respectively, act on the Fock space as:
\begin{align*}
  c_{a\sigma}\ket{\vec{n}} &=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};\\
    c^{+}_{a\sigma}\ket{\vec{n}} &=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
\end{align*}
where $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$ accounts for the fermionic sign imposed by the Pauli principle.
The bosonic operators $b$ and $b^\dagger$ act on the phonon part of $\ket{p}$ of the Fock state as:
\begin{align*}
  b\ket{p} &= \sqrt{p-1}\ket{p} \\
  b^\dagger\ket{p} &= \sqrt{p}\ket{p} \\
\end{align*}
The implementation of Fock states and operators can be found in 
{\tt ED\_AUX\_FUNX}. In particular, in this module we define the bitwise action 
of fermionic creation and annihilation operators as functions 
{\tt CDG} and {\tt C}, used throughout the code. Additionally, 
we provide the function {\tt bdecomp} for reconstructing the Fock  state bit sequence from the integer representation.

@@

\begin{lstlisting}[style=fstyle,numbers=none]
 !Binary decomposition of the ${\color{comment-color} i\in\NNN}$ with Ntot-bits
 function Bdecomp(i,Ntot) result(ivec)
    integer :: Ntot,ivec(Ntot),l,i
    logical :: busy
    do l=0,Ntot-1
       busy=btest(i,l)
       ivec(l+1)=0
       if(busy)ivec(l+1)=1
    enddo
  end function Bdecomp
\end{lstlisting}

@@

In the context of quantum impurity problems, two commonly used 
symmetries are: i) conservation of the total charge $N$ and ii) 
conservation of the total magnetization $S_z$. Although the total 
spin operator $S^2$ may also be conserved, its implementation is 
computationally challenging and typically provides only marginal 
gains, making it less practical in many cases.
As the symmetries applies to the electronic terms, here we discuss the organization of the electronic Fock states only. The presence of phonon is accounted for exploiting the tensor structure of the global Fock space.

In \NAME, three different symmetry configurations are available, 
controlled by the input variable {\tt ed\_mode} = {\bf normal}, 
{\bf superc}, or {\bf nonsu2}. A summary of the sector properties is provided in
Table~(\ref{TabSector}).

\begin{itemize}
\item{} The {\bf normal} case conserves both the electrons total occupation $N$
and total magnetization $S_z$, or equivalently, the total number
of electrons with spin up $N_\up$ and down $N_\dw$.
Optionally, the symmetry
can be extended to act independently to each orbital and spin
component, i.e.~$\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This
specific case is discussed extensively
Ref.~\cite{Amaricci2022} so we will not be further cover it here recalling that any consideration concerning the {\tt normal} case extends directly to this case as well.

@@

\item{} The {\bf nonsu2} case conserves the total
charge $N$ while allowing spin symmetry group to be not conserved.
This scenario captures effects such as local spin-orbit coupling
$\vec{L}\cdot\vec{S}$~\cite{something}, in-plane spin ordering~\cite{BellomiaKMH} or
in-plane spin-triplet exciton condensation~\cite{Amaricci2023_excitons,Blason}.  
\end{itemize}

\begin{table}%[ht]
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_\mathrm{s}}{N_\up}\binom{N_\mathrm{s}}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_\mathrm{s}-S_z-2m}\binom{N_\mathrm{s}}{N_\mathrm{s}-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv N_\up+N_\dw$ & $\binom{2N_\mathrm{s}}{N}$ \\ 
 \hline
  \label{TabSector}
\end{tabularx}
\end{center}
\caption{
  {\bf Quantum Numbers Table}.
A table summarizing the possible values of the input variable {\tt
  ed\_mode} selecting the symmetries of the problem (first column). To each value
correspond different quantum numbers listed in the second column.
The third column reports the dimension of the electronic symmetry sector $\SS_{\vec{Q}}$ for a
given value of the quantum numbers $\vec{Q}$.}
\end{table}
From a computational perspective, constructing a symmetry
sector amounts to defining a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF_e$ which relates the electronic states $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the states $\ket{I_e}$ in the Fock space.
This map is typically implemented as a rank-1
integer array whose size corresponds to the dimension of the sector $D_{\SS_{\vec{Q}}}$. 
The global Fock state $\ket{I}=\ket{I_e}\ket{p}$, including the phonon part, is univocally determined by the relation: $I\to I_e + p 2^{2 N_s}$.

@@

The presence of a symmetry reduces the electronic Hamiltonian matrix to
a block diagonal form, where each block, labeled by $\vec{Q}$, has dimension
$D_{\SS_{\vec{Q}}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix.
In the {\bf normal} case this block structure is particularly
symmetric due to the factorized nature of the sectors as discussed
above, see also Ref.~\cite{Amaricci2022}.
The analysis of the spectrum is then reduced to inspecting the
Hamiltonian within each symmetry sector, or to a subset thereof if a
additional constraints are present.

@@

\subsection{Interaction setup}\label{sSecIntSetup}
\NAME offers two distinct methods to setup the interaction terms $\hat{H}^\mathrm{int}$ in \equ{Himp}, which are controlled by the input variables {\tt
  ED\_USE\_KANAMORI} and {\tt ED\_READ\_UMATRIX}.

>>>>

\section{Implementation}\label{SecEDIpack}
In this section we introduce the generic quantum impurity problem and
present a detailed overview of the \NAME library implementation.

\subsection{The quantum impurity problem}\label{sSecQIM}
We consider a quantum impurity problem described by the Hamiltonian:
$$
\hat{H} = \hat{H}_\mathrm{imp} + \hat{H}_\mathrm{bath} + \hat{H}_\mathrm{hyb} + \hat{H}_\mathrm{ph} + \hat{H}_\mathrm{e-ph}
$$
which characterizes a multi-orbital quantum impurity
coupled to an electronic bath and local phonons (e.g. Holstein
modes). For now, we assume no specific symmetries.

\subsubsection{Impurity Hamiltonian}
The impurity Hamiltonian is split into a quadratic and an 
interacting part,
\begin{equation}\label{Himp}
  \hat{H}_\mathrm{imp}  = \hat{H}^0_\mathrm{imp} + \hat{H}^\mathrm{int}_\mathrm{imp}.
\end{equation}
The first term has the form
\begin{equation}\label{H0imp}
  \hat{H}^0_\mathrm{imp}  =
  \sum_{\a\b\sigma\sigma'}h^{0}_{\a\b\sigma\sigma'}d^\dagger_{\a\sigma}d_{\b\sigma'}
\end{equation}
and represents the quadratic part, with $d_{\a\sigma}$ ($d^\dagger_{\a\sigma}$) being the annihilation (creation)
operators for impurity electrons in 
orbital $\alpha=1,\dots,N_\a$, with the number of orbitals $N_\a$,
and spin $\sigma=\up,\dw$.
The internal structure of the non-interacting part is captured by the 
matrix $h^{0}_{\a\b\sigma\sigma'}$, which can include 
orbital-dependent on-site energies or amplitudes, spin-orbit coupling or other single-particle terms.

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can in principle contain any 
set of two-body operators: 
\begin{equation}
\hat{H}^\mathrm{int}_\mathrm{imp} = \frac{1}{2} \sum_{ijkl} d^{\dagger}_{i}d^{\dagger}_{j}U_{ijkl}d_{l}d_{k}.\label{HintUmat}
\end{equation}
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction~\cite{Georges2013ACMP}:
\begin{equation}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^\dagger_{\a\uparrow}d_{\a\downarrow}d^\dagger_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^\dagger_{\a\uparrow}d^\dagger_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow},\\
\end{split}
\label{Hint}
\end{equation}
where we introduced the occupation number operator $n_{\alpha\sigma}=d^\dagger_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Georges2013ACMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and
the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ACMP}. Different choices, preserving part
of the combined symmetry group, can be made for other numbers of orbitals~\cite{Georges2013ACMP}.

\subsubsection{Bath and Hybrization}
The coupling between the impurity and the bath is described by
\begin{equation}\label{Hbath}
  \begin{split}
    \hat{H}_\mathrm{bath} &=
    \sum_p\sum_{\a\b\sigma\sigma'}h^p_{\a\b\sigma\sigma'}a^\dagger_{p\a\sigma}a_{p\b\sigma'},\\
    %
    \hat{H}_\mathrm{hyb} &= \sum_p\sum_{\a\b\sigma\sigma'}V^p_{\a\b\sigma\sigma'}d^\dagger_{\a\sigma}a_{p\b\sigma'}+H.c., \\
\end{split}
\end{equation}
where $p=1,\dots,N_\mathrm{bath}$ indexes the
bath elements. The operators $a_{p\alpha\sigma}$
($a^\dagger_{p\alpha\sigma}$) correspond to the destruction (creation) of
bath electrons with index $p$, orbital $\alpha$ and spin $\sigma$.
Any bath element can be composed of several electronic levels
according to the chosen bath topology (see \secu{sSecBath}). The  properties of each bath element are described by the
matrix $h^p_{\a\b\sigma\sigma'}$ while the amplitude
$V^p_{\a\b\s\sigma'}$ describes the coupling with the impurity.

\subsubsection{Electron-Phonon coupling}
Finally, we include a local electron-phonon coupling on the impurity site
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_m \omega_{0m} b_m^\dagger b_m,   \\
    % 
    \hat{H}_\mathrm{e-ph} &= \sum_m [ \hat{O}_m - A_m](b_m+b_m^\dagger), \\
\end{split}
\end{equation}
where $m=1,\dots,M$ indexes the number of local phonon modes, and $b_m$
($b_m^\dagger$) are the destruction (creation) operators for a phonon with
frequency $\omega_{0m}$. Each phonon mode is coupled to a bilinear fermionic operator $\hat{O}_m= \sum_{\alpha \beta, \sigma }g^{m}_{\alpha \beta} d^\dagger_{\alpha, \sigma} d_{\beta, \sigma}$ on the impurity, while $A_m$ is a constant acting as a displacement field.  %Indeed, if the expectation value of the bosonic displacement operator $\langle b_q^\dagger + b_q\rangle$ is non-zero, a change of variables from the bosons $b^{(\dagger)}_q$ leads to new effective bosons $\tilde{b}^{(\dagger)}_q = b^{(\dagger)}_q- A_q$ for which the displacement operator has a smaller expectation value (ideally zero if we can impose $A_m = \langle \hat{O}_m\rangle$). 
For single-orbital impurities, the relevant case is $\hat{O}_1 = g_1 n_d = g_1 \sum_{\sigma} d^{\dagger}_{\sigma}d_{\sigma}$, which defines the Holstein model\cite{Holstein1959APNY}. For two-orbital models we included also the two Jahn-Teller modes\cite{Capone2010ACMP}, for which $\hat{O}_2 = g_2(n_1 - n_2)$\cite{Scazzola2023PRB} and $\hat{O}_3 = g_3 \sum_{\sigma} (d^{\dagger}_{1\sigma}d_{2\sigma}+ h.c.)$. For each phonon mode, one has to introduce a cut-off in the phonon number, which has to be chosen according to the value of the electron-phonon coupling\cite{Capone2003PRL,Capone2006PRB} and its interplay with the electron-electron interaction\cite{Capone2004PRL_2,Sangiovanni2005PRB,Paci2005PRL,Sangiovanni2006PRL,Sangiovanni2006PRB_ph,Paci2006PRB}\fixme{Broken ref}. The larger the number of excited phonons in the ground state, the larger the cutoff. For practical calculations around 20 phonons are sufficient to achieve converged results except for the deep polaronic regime\cite{Capone2003PRL}.
Choosing properly the shift $A_m$, one can reduce the number of excited phonons in the ground state and consequently the required cutoff. 
Although feasible, dealing with more than one phonon mode quickly becomes computationally very demanding, thus we will consider only one mode and we drop the $m$ index in the rest of this work.

The phonon frequency is passed to the \NAME program via the variable { \tt W0\_PH } to be read from the input file. The displacement field acting on the phonon is passed through the variable {\tt A\_PH} from the input file. For every phonon mode, the $g$ coupling can be written as a matrix $g_{\alpha \beta}$ which can be passed in two different ways. If the matrix is diagonal ($g_{\alpha \beta} = g_\alpha \delta_{\alpha \beta}$), from the input file one can simply write in order the values $g_\alpha$ in the variable {\tt G\_PH}. For any general shape one can pass to the input file a {\tt GPHfile } variable containing the name of the file storing the $g_{\alpha \beta}$ array in Fortran style.\fixme{I don't seem to understand the last two sentences at all\ldots}
If {\tt GPHfile=NONE}, the matrix $g_{\alpha\beta}$ is read from the variable {\tt G\_PH}, otherwise it is read from the {\tt GPHfile}.

\subsubsection{System setup}
We consider a quantum impurity system comprised by a single multi-orbital impurity, a bath discretized into a number $N_\mathrm{bath}$ of {\it elements} possibly endowed with an internal structure, and a number of available phonons. 
The total size of the system is determined by three contributions: i) the number of impurity orbitals $N_\a$, ii) the total count of electronic bath levels $N_\mathrm{b}$, and iii) by the maximum  number of allowed phonons $N_\mathrm{ph}$, introducing a cutoff to the unbounded dimension of the local phonon Hilbert space.
The number of bath degrees of freedom $N_\mathrm{b}$ is a
function of the bath topology and of $N_\mathrm{bath}$, which is determined internally by \NAME.
For instance, for a simple case of a purely electronic problem, each bath
element corresponds to an independent electronic level coupled to the
impurity, thus $N_\mathrm{b}\equiv N_\mathrm{bath}$. 
The total number of electronic levels in quantum impurity systems is indicated with $N_\mathrm{s}$.

The setup of the quantum impurity problem is implemented in {\tt
  ED\_SETUP} through the input variables {\tt Nspin}$=N_\sigma$,
{\tt Norb}$=N_\a$ and {\tt Nbath}$=N_\mathrm{bath}$ globally defined in {\tt
  ED\_INPUT\_VARS}. The variable {\tt Ns}$=N_\mathrm{s}$, evaluated in
{\tt ed\_setup\_dimensions}, corresponds to the total number of
electronic levels. 
The local non-interacting Hamiltonian
$h^0_{\a\b\s\s'}$ is specified using the function {\tt ed\_set\_hloc} from {\tt ED\_AUX\_FUNX}.
The setup of the bath matrices $h^p_{\a\b\s\s'}$
requires a more involved procedure, which will be illustrated in
\secu{sSecBath}.

@@

The electronic part of the Fock state $\ket{\vec{n}}$ is represented
as a binary string of length $2N_\mathrm{s}$. Thus, any such state can be encoded
in a computer using a sequence of $2N_\mathrm{s}$ bits, or equivalently, as an integer $I=0,\dots 2^{2N_\mathrm{s}}-1$ such that $\ket{\vec{n}}=\ket{I}$.  
The electronic destruction and creation operators, $c_{a\sigma}$ and $c^\dagger_{a\sigma}$ 
respectively, act on the Fock space as:
\begin{align*}
  c_{a\sigma}\ket{\vec{n}} &=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};\\
    c^\dagger_{a\sigma}\ket{\vec{n}} &=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
\end{align*}
where $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$ accounts for the fermionic sign imposed by the Pauli principle.
The bosonic operators $b$ and $b^\dagger$ act on the phonon part $\ket{p}$ of the Fock state as
\begin{align*}
  b\ket{p} &= \sqrt{p-1}\ket{p}, \\
  b^\dagger\ket{p} &= \sqrt{p}\ket{p}. \\
\end{align*}
The implementation of Fock states and operators can be found in 
{\tt ED\_AUX\_FUNX}. In particular, in that module we define the bitwise action 
of the fermionic creation and annihilation operators as functions 
{\tt CDG} and {\tt C}, used throughout the code. Additionally, 
we provide the function {\tt bdecomp} for reconstructing the Fock  state bit sequence from the integer representation.

@@

\begin{lstlisting}[style=fstyle,numbers=none]
 !Binary decomposition of the ${\color{comment-color} i\in\NNN}$ with Ntot bits
 function Bdecomp(i,Ntot) result(ivec)
    integer :: Ntot,ivec(Ntot),l,i
    logical :: busy
    do l=0,Ntot-1
       busy=btest(i,l)
       ivec(l+1)=0
       if(busy)ivec(l+1)=1
    enddo
  end function Bdecomp
\end{lstlisting}

@@

In the context of quantum impurity problems, two commonly used 
symmetries are: i) conservation of the total charge $N$ and ii) 
conservation of the total magnetization $S_z$. Although the total 
spin operator $S^2$ may also be conserved, its implementation is 
computationally challenging and typically provides only marginal 
gains, making it less practical in many cases.
As the symmetries apply to the electronic terms, here we discuss the organization of the electronic Fock states only. The presence of phonons is accounted for by exploiting the tensor structure of the global Fock space.

In \NAME, three different symmetry configurations 
controlled by the input variable {\tt ed\_mode} = {\bf normal}, 
{\bf superc}, {\bf nonsu2} are available. A summary of the sector properties is provided in
Table~(\ref{TabSector}).

\begin{itemize}
\item{} The {\bf normal} case conserves both the electrons' total occupation $N$
and total magnetization $S_z$, or equivalently, the total number
of electrons with spin up $N_\up$ and down $N_\dw$.
Optionally, the symmetry
can be extended to act independently on each orbital and spin
component, i.e.~$\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This
specific case is discussed extensively
Ref.~\cite{Amaricci2022}, so we will cover it no further here recalling that any consideration concerning the {\tt normal} case extends directly to this case as well.

@@

\item{} The {\bf nonsu2} case conserves the total
charge $N$ while allowing the spin symmetry group to be broken.
This scenario captures effects such as local spin-orbit coupling
$\vec{L}\cdot\vec{S}$~\cite{something}\fixme{Missing citation}, in-plane spin ordering~\cite{BellomiaKMH} or
in-plane spin-triplet exciton condensation~\cite{Amaricci2023_excitons,Blason}.  
\end{itemize}

\begin{table}%[ht]
\begin{tabularx}{\linewidth}{ |X|X|X| }
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\Leftrightarrow[N_\up,N_\dw]$ &
                                                $\binom{N_\mathrm{s}}{N_\up}\binom{N_\mathrm{s}}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_\mathrm{s}-S_z-2m}\binom{N_\mathrm{s}}{N_\mathrm{s}-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv N_\up+N_\dw$ & $\binom{2N_\mathrm{s}}{N}$ \\ 
 \hline  
\end{tabularx}
\vspace{2mm}
\caption{\label{TabSector}
  {\bf Quantum Numbers}.
A table summarizing the possible values of the input variable {\tt
  ed\_mode} selecting the symmetries of the problem (first column). To each value
correspond different quantum numbers listed in the second column.
The third column reports the dimension of the electronic symmetry sector $\SS_{\vec{Q}}$ for a
given value of the quantum numbers $\vec{Q}$.}
\end{table}

From a computational perspective, constructing a symmetry
sector amounts to defining an injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF_e$ that relates the electronic states $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the states $\ket{I_e}$ in the Fock space.
This map is typically implemented as a rank-1
integer array, whose size corresponds to the dimension of the sector $D_{\SS_{\vec{Q}}}$. 
The global Fock state $\ket{I}=\ket{I_e}\ket{p}$, including the phonon part, is univocally determined by the relation: $I\to I_e + p 2^{2 N_s}$.

@@

The presence of a symmetry reduces the electronic Hamiltonian matrix to
a block-diagonal form, where each block labeled by $\vec{Q}$ has dimension
$D_{\SS_{\vec{Q}}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix.
In the {\bf normal} case this block structure is particularly
symmetric due to the factorized nature of the sectors as discussed
above, see also Ref.~\cite{Amaricci2022}.
The analysis of the spectrum is then reduced to inspecting the
Hamiltonian within each symmetry sector, or to a subset thereof if
additional constraints are present.

@@

\subsection{Interaction setup}\label{sSecIntSetup}
\NAME offers two distinct methods to setup the interaction terms $\hat{H}^\mathrm{int}_\mathrm{imp}$ in \equ{Himp}, which are controlled by the input variables {\tt
  ED\_USE\_KANAMORI} and {\tt ED\_READ\_UMATRIX}.

The generalized Hubbard-Kanamori interaction, as defined in \equ{Hint}, is natively
supported and controlled by the input parameters {\tt ULOC}$=U$, {\tt
  UST}$=U'$, {\tt JH}$=J$, {\tt JX}$=J_{x}$ and {\tt JP}$=J_{p}$.
These quantities can be specified either in the input file or directly
via the command line, provided the logical input parameter {\tt ED\_USE\_KANAMORI=T} and setting {\tt ED\_READ\_UMATRIX=F}. In this mode, the maximum number of impurity orbitals {\tt NORB}$=N_\a$ is limited to $5$.
